---
title: "Hotel booking Analysis"
author: "*Buthaina Alshareef*"
date: "*21-10-2020*"
output:
  html_document:
   toc: true
   toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(DT)
#library(Rmisc)
```



```{r get familiar with data, echo=FALSE, message=FALSE}
Hotel_bookings <- read_csv("hotel_bookings.csv")

 summary(Hotel_bookings)
# names(Hotel_bookings)
# glimpse(Hotel_bookings)
# typeof(Hotel_bookings)
# str(Hotel_bookings)
# class(Hotel_bookings)

```
# Descriptive Statistics

```{r, echo=FALSE, message=FALSE}
#range(Hotel_bookings$adr) (0, 6000)
#range(Hotel_bookings$stays_in_week_nights)
HotelTable <- Hotel_bookings %>% 
            group_by(hotel) %>%
             summarise(n = nrow(Hotel_bookings),
            Average_days = mean(stays_in_week_nights),
            SD_days = sd(stays_in_week_nights),
            min_days = min(stays_in_week_nights),
            max_days = max(stays_in_week_nights),
            Average_price = mean(adr),
            SD_price = sd(adr),
            min_price = min(adr),
            max_price = max(adr)
            )


datatable(HotelTable)
```

**Days Plot**

```{r ,echo=FALSE, message=FALSE}
ggplot(Hotel_bookings, aes(x = stays_in_week_nights, color = hotel)) +
   geom_histogram(bins = 9) +
  geom_vline(aes(xintercept=mean(stays_in_week_nights)),
            color="blue", linetype="dashed", size=1) +
  coord_cartesian(xlim = c(0,50)) +
  facet_grid(. ~ hotel) +
  labs(title = "mean & frequency for the days") +
  theme_classic()+
 NULL
```

**Price Plot**

```{r}
ggplot(Hotel_bookings, aes(x = adr, color = hotel)) +
   geom_histogram(bins = 9) +
  geom_vline(aes(xintercept=mean(adr)),
            color="blue", linetype="dashed", size=1)+
  #coord_cartesian(xlim = c(0,6000)) +
  facet_grid(. ~ hotel) +
  labs(title = "mean & frequency for the price") +
  theme_classic() +
 NULL
```

Cont with Cont 
Adr 
sty week nights 
Sample 






## Cat ~ Cat 
Hotel is binary variable 
is_canceled is binary variable 

**the total of canceled reservations to not canceled by hotel type (proportion)**
Bar chart 
```{r, echo=FALSE, message=FALSE}
ggplot(Hotel_bookings, aes(x = hotel,
         y = prop.table(stat(count)),
         fill = factor(is_canceled),
       label = percent(prop.table(stat(count))))) +
  geom_bar( position = "dodge") +
  geom_text(
    stat = "count",
    position = position_dodge(.9),
    vjust = -0.5,
    size = 3
  ) +
 scale_y_continuous(labels = percent) +
  labs(x = "Hotel Type", y = "Proportion", title = "Cancelation Status by Hotel Type") +
  scale_fill_discrete(name = "Booking Status",
                      breaks = c("0","1"),
                      labels = c("Cancelled", "Not Cancelled")) +
  theme_classic() +
NULL
```











## Two sample t-test

Categorical described by Continuous
Are the two types of hotel have the same prices? 


```{r}
ggplot(Hotel_bookings, aes(x = 1, y = adr)) +
  geom_jitter(width = 0.2) +
  labs(title = "No information from X at all")
```

```{r}
ggplot(Hotel_bookings, aes(x = hotel, y = adr)) +
  geom_jitter(width = 0.2) +
  labs(title = "ALL information from X that is available")
```
**two sample t.test**
$\bar{x}_{City Hotel} - \bar{x}_{Resort Hotel}$
```{r}
t.test(adr ~ hotel, data = Hotel_bookings)
```
Because the p-value < 2.2e-16, we are going to:
- Reject the null hypothesis 
- There is evidence that the hotel type is related to the price 
- There is an association between price and hotel type
- We can use the hotel type to predict price!

# 95% CI

```{r, echo=FALSE, message=FALSE}
# Data
city_hotel <- Hotel_bookings$stays_in_week_nights[Hotel_bookings$hotel == "City Hotel"]
resort_hotel <- Hotel_bookings$stays_in_week_nights[Hotel_bookings$hotel == "Resort Hotel"]
```


```{r, echo=FALSE, message=FALSE}
# city_hotel_CI <- CI(city_hotel, ci = 0.95)
# resort_hotel_CI <- CI(resort_hotel, ci = 0.95)
```









# Linear Regression
**Does the hotel price affect the number of staying days in weekdays?**
```{r}
City_price_mean <- mean(Hotel_bookings$adr[Hotel_bookings$hotel == "City Hotel"])
Hotel_bookings %>% 
  filter(hotel == "City Hotel") %>% 
  ggplot(aes(x = 1, y = adr)) +
  geom_jitter(width = 0.2) +
  geom_hline(aes(yintercept = City_price_mean, color = "Null model (y-bar)")) +
  coord_cartesian(xlim = c(0,2)) +
  labs(title = "No information from X at all")
```

```{r}
posn_j <- position_jitter(seed = 136)

Hotel_bookings %>% 
  filter(hotel == "City Hotel") %>% 
ggplot(aes(x = stays_in_week_nights, y = adr)) +
  geom_point(shape = 16, alpha = 0.25) +
  # stat_summary(fun.data = mean_sdl, 
               #fun.args = list(mult = 1), col = "red") +
  geom_smooth(method = "lm", se = FALSE) +
  NULL
```
- reject the null hypothesis  
- There is association between price and the number of stay days 

```{r}
speed_eye <- lm(adr ~ stays_in_week_nights, data = Hotel_bookings[Hotel_bookings$hotel == "City Hotel",])
summary(speed_eye)
```
Prove it by p-value: < 2.2e-16 

```{r}
Hotel_bookings %>%
  sample_frac(0.05) %>%
  ggplot( aes(x=1, y=adr)) +
    geom_point(color="#69b3a2", shape = 1, position = posn_j ) +
  geom_hline(aes(yintercept = City_price_mean, color = "Null model (y-bar)")) +
  NULL
```


```{r}
Hotel_bookings %>%
  sample_frac(0.05) %>%
  ggplot( aes(x=stays_in_week_nights, y=adr, color = hotel)) +
    geom_point(shape = 1, position = posn_j) +
  geom_smooth(method = "lm", se = FALSE) +
  NULL
```
