---
title: "Student_performance"
author: "*Buthaina Alshareef*"
date: "*21-10-2020*"
output:
  html_document:
   toc: true
   toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(DT)
library(ggpubr)
library(knitr)

```



```{r get familiar with data, echo=FALSE, message=FALSE}
# https://www.kaggle.com/spscientist/students-performance-in-exams
student <- read.csv("StudentsPerformance.csv")

# names(student) <- make.names(names(student), unique = TRUE)
posn_j <- position_jitter(seed = 136)
#summary(student)
names(student)
# glimpse(student)
# typeof(student)
# str(student)
#class(student)
#describe(student)

```


# Descriptive Statistics


```{r, echo=FALSE, message=FALSE}
#range(student$reading.score) (0, 6000)
#range(student$math.score)
studentTable1 <- student %>% 
  group_by(gender) %>%
  summarise(n = n(),
            Average_math = mean(math.score),
            SD_math = sd(math.score),
            min_math = min(math.score),
            max_math = max(math.score))

studentTable2 <- student %>% 
  group_by(gender) %>%
  summarise(n = nrow(student),
            Average_reading = mean(reading.score),
            SD_reading = sd(reading.score),
            min_reading = min(reading.score),
            max_reading = max(reading.score))

studentTable3 <- student %>% 
  group_by(gender) %>%
  summarise(n = nrow(student),
            Average_writing = mean(writing.score),
            SD_writing = sd(writing.score),
            min_writing = min(writing.score),
            max_writing = max(writing.score))



student_gender_summary <- student %>% 
  mutate(ID = 1:nrow(student)) %>% 
  pivot_longer(ends_with("score"), names_to = "test", values_to = "score") %>% 
  mutate(test = str_remove(test, ".score")) %>% 
  group_by(gender, test) %>%
  summarise(n = n(),
            Average = mean(score),
            SD = sd(score),
            min = min(score),
            max = max(score))

```

### Summary of scores by gender

```{r , echo=FALSE, message=FALSE}
kable(student_gender_summary)
```

```{r}
ggplot(student_gender_summary, aes(test, Average, ymin = Average - SD, ymax = Average + SD, color = gender)) +
  geom_pointrange(position = position_dodge(0.3))
```



## Inferential statistics

#### Normal Distribution

```{r}
ggplot(student, aes(x = math.score, fill = gender, color = gender)) +
  geom_density(alpha = 0.2) +
  geom_vline(aes(xintercept = Average, color = gender), data = student_gender_summary[student_gender_summary$test == "math",] ,
             linetype = "dashed", size = 1) +
  coord_cartesian(xlim = c(0,100))
```

# how to get all three score distributions together

```{r}
student %>% 
  mutate(ID = 1:nrow(student)) %>% 
  pivot_longer(ends_with("score"), names_to = "test", values_to = "score") %>% 
  mutate(test = str_remove(test, ".score")) %>% 
  select(gender, test, score) %>% 
  ggplot(aes(x = score, fill = gender, color = gender)) +
  geom_density(alpha = 0.2) +
  geom_vline(aes(xintercept = Average, color = gender), 
             data = student_gender_summary,
             linetype = "dashed", size = 1) +
  facet_grid(test ~ .) +
  coord_cartesian(xlim = c(0,100))
```




```{r ,echo=FALSE, message=FALSE}
ggplot(student, aes(x = math.score)) +
  geom_histogram(aes(y=..density..),
                 fill = "#deebf7") +
  geom_density(alpha=0.2, fill = "#FF6666", color = "#3182bd") +
  geom_vline(aes(xintercept = mean(math.score)), linetype = "dashed", size = 1, color = "#de2d26") +
  facet_grid(. ~ gender) +
  coord_cartesian(xlim = c(0,100)) +
  labs(title = "Population distribution for the math score", x = "math score",y = "Frequency") +
  # scale_color_brewer("Iris Species", palette = "Dark2") +
  theme_classic()+
  NULL

```

### Sample means for Math Score

```{r,echo=FALSE, message=FALSE}
n <- 3
r <- 30000
# Take r samples, each having n observations:
math_sample_means <- rep(n, r) %>% 
  map_dbl( ~  mean(sample(student$math.score, size = ., replace = TRUE)) )
# plot them:
h <- hist(math_sample_means, breaks = 100, col = "#9ecae1", xlab = " math score" , main = "Central Limit theorem")
xfit <- seq(min(math_sample_means), max(math_sample_means), length = 40) 
yfit <- dnorm(xfit, mean = mean(math_sample_means), sd = sd(math_sample_means)) 
yfit <- yfit * diff(h$mids[1:2]) * length(math_sample_means) 

lines(xfit, yfit, col = "#de2d26", lwd = 2)
```


```{r ,echo=FALSE, message=FALSE}
ggplot(student, aes(x = reading.score)) +
  geom_histogram(aes(y=..density..), alpha=0.5, 
                 position="identity", color = "#3182bd", fill = "#deebf7") +
  geom_density(alpha=0.2, fill = "#FF6666", color = "#3182bd") +
  geom_vline(aes(xintercept=mean(reading.score)),
             linetype="dashed", size=1, color = "#de2d26") +
  #facet_grid(. ~ gender) +
  coord_cartesian(xlim = c(0,100)) +
  labs(title = "Population distribution for the reading score", x = "reading score", y = "Frequency") +
  theme_classic()+
  NULL

```

### Sample means for Reading Score
```{r,echo=FALSE, message=FALSE}
n <- 3
r <- 30000
# Take r samples, each having n observations:
reading_sample_means <- rep(n, r) %>% 
  map_dbl( ~  mean(sample(student$reading.score, size = ., replace = TRUE)) )
# plot them:
h <- hist(reading_sample_means, breaks = 100, col = "#9ecae1", xlab = "reading score" , main = "Central Limit theorem")
xfit <- seq(min(reading_sample_means), max(reading_sample_means), length = 40) 
yfit <- dnorm(xfit, mean = mean(reading_sample_means), sd = sd(reading_sample_means)) 
yfit <- yfit * diff(h$mids[1:2]) * length(reading_sample_means) 

lines(xfit, yfit, col = "#de2d26", lwd = 2)
```


```{r ,echo=FALSE, message=FALSE}
ggplot(student, aes(x = writing.score)) +
  geom_histogram(aes(y=..density..), alpha=0.5, 
                 position="identity", color = "#3182bd", fill = "#deebf7") +
  geom_density(alpha=0.2, fill = "#FF6666", color = "#3182bd") +
  geom_vline(aes(xintercept=mean(writing.score))
             , linetype="dashed", size=1, color = "#de2d26") +
  #facet_grid(. ~ gender) +
  coord_cartesian(xlim = c(0,100)) +
  labs(title = "Population distribution for the writing score", x = "writing score",y = "Frequency") +
  theme_classic()+
  NULL

```

### Sample means for Writing Score
```{r,echo=FALSE, message=FALSE}
n <- 3
r <- 30000
# Take r samples, each having n observations:
writing_sample_means <- rep(n, r) %>% 
  map_dbl( ~  mean(sample(student$writing.score, size = ., replace = TRUE)) )
# plot them:
h <- hist(writing_sample_means, breaks = 100, col = "#9ecae1", xlab = "writing score" , main = "Central Limit theorem")
xfit <- seq(min(writing_sample_means), max(writing_sample_means), length = 40) 
yfit <- dnorm(xfit, mean = mean(writing_sample_means), sd = sd(writing_sample_means)) 
yfit <- yfit * diff(h$mids[1:2]) * length(writing_sample_means) 

lines(xfit, yfit, col = "#de2d26", lwd = 2)
```

- The data is normally distributed in math, reading, and writing score in both population and sample means 
-  the sample mean $\bar{x}$ is an unbiased estimator for the underlying population mean $\mu$
- the sample means $\bar{x}$ are centered on the population mean  $\mu$



# 95% CI

Math score
```{r,echo=FALSE, message=FALSE}
math_mu <- mean(student$math.score)
t.test(student$math.score, mu = math_mu, paired = FALSE )
```
There is a 95% chance that interval [65.14806, 67.02994] _covers_ the true population parameter $\mu$ (66.089).

```{r,echo=FALSE, message=FALSE}
reading_mu <- mean(student$reading.score)
t.test(student$reading.score, mu = reading_mu, paired = FALSE )
```
There is a 95% chance that interval [68.26299, 70.07501] _covers_ the true population parameter $\mu$ (69.169).
```{r,echo=FALSE, message=FALSE}
writing_mu <- mean(student$writing.score)
t.test(student$writing.score, mu = writing_mu, paired = FALSE )
```
There is a 95% chance that interval [67.11104, 68.99696] _covers_ the true population parameter $\mu$ (68.054).

# Probability 
```{r,echo=FALSE, message=FALSE}
math_50 <-  sum(student$math.score > 50)
n = nrow(student)
math_50_prob <- (math_50/n)*100
```
The probability for the students to get more than 50 in math is `r math_50_prob`%

```{r,echo=FALSE, message=FALSE}
writing_50 <-  sum(student$writing.score > 50)
n = nrow(student)
writing_50_prob <- (writing_50/n)*100
```
The probability for the students to get more than 50 in writing is `r writing_50_prob`%

```{r,echo=FALSE, message=FALSE}
reading_50 <-  sum(student$reading.score > 50)
n = nrow(student)
reading_50_prob <- (reading_50/n)*100
```
The probability for the students to get more than 50 in reading is `r reading_50_prob`%

# Two sample t-test

Is there a relationship between gender and scores in each subject?
```{r,echo=FALSE, message=FALSE}
ggplot(student, aes(x = 1, y = math.score)) +
  geom_point(position = posn_j, shape = 16, color = "#3182bd") +
  labs(title = "No information from X at all", y = "math score") +
  theme_classic() +
  NULL
```

```{r,echo=FALSE, message=FALSE}
ggplot(student, aes(x = gender, y = math.score, color = gender)) +
  geom_point(position = posn_j, shape = 16) +
  scale_color_manual(values=c("#fa9fb5", "#3182bd")) +
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1), col = "#de2d26") +
  labs(title = "ALL information from X that is available", y = "math score") +
  NULL
```
$\bar{x}_{female} - \bar{x}_{male}$
```{r,echo=FALSE, message=FALSE}
t.test(math.score ~ gender, data = student)

```

Here the p-value =  8.421e-08 which means: 

- Reject the null hypothesis 
- There is evidence that the gender is related to the scores in math 
- We can use the gender to predict scores in math!

```{r,echo=FALSE, message=FALSE}
#t.test(math.score ~ gender, data = student, alternative = "less")
#Test the normality
# with(student, shapiro.test(math.score[gender == "male"]))
```

```{r,echo=FALSE, message=FALSE}
# with(student, shapiro.test(math.score[gender == "female"]))
```

# Linear regression 

Is there a relationship between math score and scores in each subject?

```{r,echo=FALSE, message=FALSE}
ggplot(student, aes(x = math.score, y = reading.score)) +
  geom_point(shape = 16, position = posn_j, color = "#3182bd" ) +
  geom_smooth(method = "lm", se = FALSE, color = "#de2d26" ) +
  labs(x = "math score", y = "reading score") +
  theme_classic() +
  NULL
```

t-test
```{r,echo=FALSE, message=FALSE}
math_reading <- lm(reading.score ~ math.score, data = student)
summary(math_reading)
```

Reject the null hypothesis 
There is a relashion between math score and reading score 

```{r,echo=FALSE, message=FALSE}
ggplot(student, aes(x = math.score, y = writing.score)) +
  geom_point(shape = 16, position = posn_j, color = "#3182bd") +
  geom_smooth(method = "lm", se = FALSE, color = "#de2d26") +
  labs(x = "math score", y = "writing score") +
  theme_classic() +
  NULL
```

t-test

```{r,echo=FALSE, message=FALSE}
math_writing <- lm(writing.score ~ math.score, data = student)
summary(math_writing)
```
Reject the null hypothesis 

# One-way ANOVA

Visualizing the relashionship between each group of education 
```{r,echo=FALSE, message=FALSE}

#unique(student$parental.level.of.education)
ggline(student, x = "parental.level.of.education", y = "math.score", 
       add = c("mean_se", "jitter"), 
       order = c("high school", "some high school", "some college", "bachelor's degree", "master's degree", "associate's degree"),
       ylab = "math score", 
       xlab = "education", 
       point.color = "#de2d26", 
       title = "Parental level of education",
       color = "#3182bd"
)
```

The lowest scores illustrates that the students whose parent have high school education 
```{r,echo=FALSE, message=FALSE}
res.aov <- aov(math.score ~ parental.level.of.education, data = student)
summary(res.aov)
```
the low p-value indicates that the parental level of education means are different, thus the null hypothesis would be rejected.

```{r,echo=FALSE, message=FALSE}
tukeyTest <- TukeyHSD(res.aov)

datatable(tukeyTest$parental.level.of.education, colnames = c("Group Comparison", "Difference in Means", "Upper 95% CI", "Lower 95% Cl", "Adjusted p-value"))
```

